### 2024-12-14 analysis.qmd
Update code

---
title: Plots and data summaries for manuscript
author: Haziq Jamil
page-layout: full
grid:
  sidebar-width: 0px
editor_options: 
  chunk_output_type: console
html-table-processing: none
---

## Load data

```{r}
#| echo: true
#| eval: true 
#| file: data.R
```

## Summary of variables

```{r}
#| label: tbl-summary
#| tbl-cap: Summary of housing data
hsp |>
  mutate(
    price = price / 1000,
    storeys = factor(
      ifelse(storeys >= 3, "3+", as.character(storeys)),
      levels = c("1", "2", "3+")
    )
  ) |>
  tbl_summary(
    include = c(price:baths),
    by = district,
    missing = "no", 
    type = all_continuous() ~ "continuous2",
    statistic = list(
      all_continuous2() ~ c("{mean} ({sd})", Range = "{min} - {max}", "{median} ({p25}, {p75})")
    ),
    label = list(
      price = "Price (BND 1,000)",
      type = "Property type",
      plot_area = "Plot area (acres)",
      floor_area = "Floor area (sq. ft.)",
      storeys = "Number of storeys",
      status = "Development status",
      tenure = "Land tenure",    
      beds = "Number of bedrooms",
      baths = "Number of bathrooms"
    ),
    digits = list(
      baths = 1,
      beds = 1
    )
  ) |>
  add_overall() |>
  add_n() |>
  modify_header(label ~ "**Variable**")
```

## Correlations

```{r}
#| label: fig-corr
#| fig-cap: Pairwise correlation plot of continuous variables.
#| warning: false
#| message: false
#| fig-width: 7.5
#| fig-height: 7.5
#| out-width: 100%

my_fn <- function(data, mapping, method = "lm", ...) {
  ggplot(data = data, mapping = mapping) + 
    geom_point(alpha = 0.5) + 
    geom_smooth(method = method, formula = y ~ x, se = FALSE,
                col = RColorBrewer::brewer.pal(3, "Set1")[2], ...)
}

pm <-
  hsp |>
  mutate(
    logprice = log(price),
    price = price / 1000
  ) |>
  select(
    `Plot area` = plot_area,
    `Floor area` = floor_area,
    Beds = beds,
    Baths = baths,
    `Price (BND 1,000)` = price,
    `Log Price` = logprice
  ) |>
  ggpairs(
    progress = FALSE,
    lower = list(continuous = my_fn)
  ) +
  theme_bw() +
  scale_x_continuous(labels = scales::number) +
  scale_y_continuous(labels = scales::number) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
pm[6,5] <- pm[5, 6] <- NULL
pm
```

## Data availability by year

```{r}
#| label: tbl-avail
#| tbl-cap: Data availability by year
hsp |>
  mutate(
    year = year(date),
    missing_price = is.na(price),
    missing_type = is.na(type),
    missing_tenure = is.na(tenure),
    missing_status = is.na(status),
    missing_plot_area = is.na(plot_area),
    missing_floor_area = is.na(floor_area),
    missing_storeys = is.na(storeys),
    missing_beds = is.na(beds),
    missing_baths = is.na(baths),
    missing_housechar = missing_plot_area & missing_floor_area & missing_beds & missing_baths,
    year = year(quarter),
  ) |>
  summarise(
    count = n(),
    spatial = length(unique(mukim)) / nrow(bruneimap::mkm_sf),
    # price = sum(missing_price),
    type = sum(missing_type),
    house_char = sum(missing_housechar),
    source = list(unique(method)),
    .by = year
  ) |>
  mutate(
    `nat-archive` = map_chr(source, ~ ifelse("nat-archive" %in% .x, "✔", "")),
    `online-archive` = map_chr(source, ~ ifelse("online-archive" %in% .x, "✔", "")),
    `web-scrape` = map_chr(source, ~ ifelse("web-scrape" %in% .x, "✔", "")),
    llm = map_chr(source, ~ ifelse("web-scrape-llm" %in% .x, "✔", ""))
  ) |> 
  select(-source) |>
  mutate(across(type:house_char, \(x) x / count)) |>
  gt(rowname_col = "year") |>
  cols_align(
    align = "center",
    columns = `nat-archive`:llm
  ) |>
  tab_spanner(
    label = "Data source",
    columns = `nat-archive`:`web-scrape`
  ) |>
  tab_spanner(
    label = "Missing data severity",
    columns = type:house_char
  ) |>
  fmt_percent(
    spatial:house_char,
    decimals = 1
  ) |>
  cols_label(
    year ~ "Year",
    count ~ "Count",
    spatial ~ "Spatial coverage (mukim)",
    type ~ "Property Type",
    house_char ~ "Property Characteristics",
    `nat-archive` ~ "National Archive",
    `online-archive` ~ "Online Archive",
    `web-scrape` ~ "Web Scraping",
    llm ~ "LLM post-processing"
  ) |>
  grand_summary_rows(
    columns = count:house_char,
    fns = list(Mean = "mean"),
    fmt = list(
      ~ fmt_number(., decimals = 0, columns = "count"),
      ~ fmt_percent(., decimals = 1, columns = c("spatial", "type", "house_char"))
    )
  ) |>
  tab_footnote(
    footnote = paste0("Total number of mukims in Brunei = ", nrow(bruneimap::mkm_sf), "."),
    locations = cells_column_labels(columns = spatial)
  ) |>
  tab_footnote(
    footnote = "Unknown property type.",
    locations = cells_column_labels(columns = type)
  ) |>
  tab_footnote(
    footnote = "Missing all of plot area, floor area, beds, and baths variables.",
    locations = cells_column_labels(columns = house_char)
  ) |>
  tab_options(
    quarto.disable_processing = TRUE,
    grand_summary_row.text_transform = "capitalize"
  ) 
```

## Spatial distribution

```{r}
#| label: fig-spatial
#| fig-cap: Spatial distribution of median property prices by mukim.
#| message: false
#| warning: false
ggplot(hsp_mkm) +
  geom_sf(aes(fill = price), col = "gray50", linewidth = 0.5) +
  scale_fill_viridis_c(
    option = "cividis",
    na.value = "transparent",
    labels = scales::dollar,
    trans = scales::pseudo_log_trans(sigma = 0.001),
    name = "Median\nprice"
  ) +
  ggrepel::geom_label_repel(
    data = drop_na(hsp_mkm, price) |> 
      mutate(mukim = gsub("Mukim ", "", mukim)),
    aes(label = mukim, geometry = geometry),
    size = 2.7,
    stat = "sf_coordinates",
    max.overlaps = Inf,
    min.segment.length = 0,       
    segment.size = 0.3,           
    segment.curvature = 0.1,      
    force = 5                     
  ) +
  labs(x = NULL, y = NULL) +
  theme_bw()
```

## Price evolution

```{r}
#| label: fig-price-evolution
#| fig-cap: "Median smoothed prices per square foot by property type using a 24-month (8-quarter) rolling window."
#| message: false
#| out-width: 100%
#| fig-height: 5
#| fig-width: 9
slider::slide_period_dfr(hsp_all, hsp_all$date, "month", \(df) {
  df |>
    filter(type != "Land") |>
    summarise(
      date = min(date), 
      price = median(price, na.rm = TRUE),
      plot_area = median(plot_area, na.rm = TRUE),
      floor_area = median(floor_area, na.rm = TRUE),
      .by = type
    )
}, .before = 18, .after = 6) |>
  distinct(date, type, .keep_all = TRUE) |> 
  mutate(
    price_per_sqft = price / floor_area,
    type = factor(type, levels = c("Detached", "Semi-Detached", "Terrace",
                                   "Apartment", "Overall")) 
  ) |>
  ggplot(aes(x = date, y = price_per_sqft, col = type)) +
  geom_line(aes(linewidth = type)) +
  scale_x_date(
    breaks = scales::breaks_width("1 year"), 
    labels = scales::label_date("%Y"),
    name = NULL
  ) +
  scale_y_continuous(
    labels = scales::dollar,
    name = "Price per square foot (BND)"
  ) +
  scale_colour_manual(values = c(RColorBrewer::brewer.pal(4, "Set1"), "black")) +
  scale_linewidth_manual(values = c(rep(0.6, 4), 1.2)) +
  labs(
    col = NULL,
    linewidth = NULL
    # caption = "Median smoothed prices using a 12-month rolling window."
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  ) 
```

## RPPI Comparison

```{r}
#| label: fig-rppi
#| fig-cap: "Comparison of quarterly median price per square foot indices (Median PPSF) and the official Residential Property Price Index (RPPI) from Brunei Darussalam Central Bank (BDCB)."
#| message: false
#| out-width: 100%
#| fig-height: 5
#| fig-width: 9

hsp_rppi |>  
  pivot_longer(c(index, rppi), names_to = "series", values_to = "value") |>
  ggplot(aes(x = quarter, y = value, col = series)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_line(linewidth = 0.8) +
  scale_colour_brewer(palette = "Set1") +
  scale_y_continuous(labels = scales::percent, name = "Index") +
  zoo::scale_x_yearqtr(
    format = "%Y-Q%q", 
    name = NULL, 
    breaks = seq(2015, 2024.5, by = 0.25)
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  geom_text(
    data = tibble(
      quarter = c(2024.3, 2024.3),
      series = c("index", "rppi"),
      value = c(0.941, 0.955),
      label = c("Median PPSF", "RPPI (BDCB)")
    ),
    aes(label = label),
    hjust = 0
  ) +
  coord_cartesian(xlim = c(2015, 2025))
```