### 2025-02-22 _submission/Revision 1/_response.qmd
Update figure llm

---
title: 'Review of "Random Item Response Data Generation Using a Limited-
Information Approach: Applications to Assessing Model
Complexity"'
subtitle: 'Response to reviewers'
date: today
format: 
  pdf:
    papersize: a4
    colorlinks: true
    citecolor: blue
    link-citations: true
bibliography: ../../refs-HJ.bib
csl: ../../data-in-brief.csl
---

## Reviewer 1

## Reviewer 2

## References